---
title: "IC50 from MTT data"
author: "Raymond Otoo"
date: "8/11/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("/Users/raymondotoo/OneDrive\ -\ University\ of\ Arkansas\ for\ Medical\ Sciences/Allen\ Lab\ works/MTT\ analysis/nalm_6")
```

```{r}
#library(xlsx)
library("readxl")
mtt_excelFile <-  read_excel("mttnalm6_48hr0810222.xlsx",sheet = "Sheet1")
head(mtt_excelFile)
colnames(mtt_excelFile)
```

```{r}
df <- data.frame (uM  = c("DMSO","Medium",0.001,0.005,0.01, 0.05,0.1,0.5,1,5,10,15,20,25,30,35,40,50),
                  Rep1 = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
                  Rep2 = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
                  Rep3 = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0),
                  Rep4 = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
                  )
df

rownames(df)
```

```{r}
#`````````````````````````````` Funtion that reorganises the plate info with a loop````````````````````````````````````````
format_mttResults <- function(df,mtt_excelFile){
  i = 1
  j = 1
  while (i< 19) {
    df[i,c(2,3,4,5)]<- mtt_excelFile[j,c(2,3,4,5)]
    df[i+1,c(2,3,4,5)]<- mtt_excelFile[j,c(6,7,8,9)]
    df[i+2,c(2,3,4,5)]<- mtt_excelFile[j,c(10,11,12,13)]
    
    #keep track of the indices for the plate layout of the dataframe and the raw mtt dataFrame.
    j=j+1
    i=i+3
  }
 return (df) 
}

format_mttResults(df,mtt_excelFile)

#````````````````````````````````````````````````````````````````````````````````````````````````````````````````````````
```

```{r}
library(dplyr, warn.conflicts = FALSE)
df<-df %>% rowwise() %>% mutate(meann = mean(c(Rep1,Rep2,Rep3,Rep4)))
df<-df %>% rowwise() %>% mutate(Std_dev = sd(c(Rep1,Rep2,Rep3,Rep4)))

#evaluate an expression in the context, of a data frame
df$variation <- with(df, (Std_dev / meann) * 100)
df
```

```{r}
#calculate variability for each of the replicates
df$Rep1_percent_var <- with(df, ((Rep1*100)/df$meann[2]))
df$Rep2_percent_var <- with(df, ((Rep2*100)/df$meann[2]))
df$Rep3_percent_var <- with(df, ((Rep3*100)/df$meann[2]))
df$Rep4_percent_var <- with(df, ((Rep4*100)/df$meann[2]))

df<-df %>% rowwise() %>% mutate(var_avg = mean(c(Rep1_percent_var,Rep2_percent_var,Rep3_percent_var,Rep4_percent_var)))
df<-df %>% rowwise() %>% mutate(var_std = sd(c(Rep1_percent_var,Rep2_percent_var,Rep3_percent_var,Rep4_percent_var)))
df$ug <- with(df,df$uM )

#replace the meduim and DMSO cells with zero(s)
df$ug[1] <- with(df,0 )
df$ug[2] <- with(df,0 )

#convert the ug character vector to numeric temporarily. Also convert  the ug to Grams
df$G <- with(df,as.numeric(df$ug) *0.000001 )

#create a column with the log10 of the concentrations of the drugs in Gram
df$LogC <- with(df,log10(df$G))
df
```

